REVISITING SEMI-CONTINUOUS HIDDEN MARKOV MODELS

### 2011-09-27 misc/papers/icassp12_tied/icassp12.tex
submitted version
git-svn-id: https://svn.code.sf.net/p/kaldi/code/sandbox/discrim@525 5e6a8d80-dfce-4ca6-a32a-6e07a63d50c8

\begin{table}%[tb]
\begin{center}
%\footnotesize
%\begin{tabular}{|l||r|r|r|r|r|r||c|}
\begin{tabular}{|l||c|c|c|c|c|c||c|}
\hline
~               & ma87 & oc87 & fe89 & oc89 & fe91 & se92 & {\em avg}  \\ \hline\hline
%{\em mono/cont} &  6.24 &  8.72 &  7.85 &  9.80 &  8.21 & 13.60 & 9.50 \\ \hline
%{\em tri1/cont} &  0.96 &  2.76 &  2.69 &  3.61 &  3.30 &  6.33 & 3.64 \\ \hline\hline
{\em cont} &  1.08 &  2.48 &  2.69 &  3.46 &  2.66 &  5.90 & 3.38 \\ \hline
{\em semi} &  1.80 &  3.19 &  4.72 &  4.62 &  4.15 &  6.88 & 4.66 \\ \hline 
{\em 2lvl} &  0.48 &  1.70 &  2.46 &  3.35 &  1.89 &  5.31 & {\bf 2.90} \\ \hline
{\em sgmm} &  0.48 &  2.20 &  2.62 &  2.50 &  1.93 &  5.12 & 2.78 \\ \hline
\end{tabular}
\end{center}
\caption{\label{tab:res_rm}
Detailed recognition results in \% WER for the six DARPA test sets using 
different acoustic models on the RM data set.
}
\end{table}

@@

% TODO Korbinian: More expts are ongoing; these include the full training set
%                 and cmvn, and hopefully a good param combination for 2lvl...
\begin{table}%[tb]
\begin{center}
\begin{tabular}{|l||c|c|c|c|c|c||c|}
\hline
~          & eval '92 & eval '93 & {\em avg} \\ \hline\hline
{\em cont} &    12.75 &    17.10 &     14.39 \\ \hline % si-84/half, cmvn

% tri2-2lvl-208-3072-4000-0-0-0-0 !! no cmvn !!
%{\em 2lvl} &  13.95 &  21.61 &     16.85 \\ \hline

% tri2-2lvl-208-4096-6000-0-1-35-0.2 !! no cmvn !!
{\em 2lvl} &    12.83 &    21.61 &     16.16 \\ \hline

{\em sgmm} &    10.76 &    17.82 &     13.44 \\ \hline % si-84/half, cmvn
\end{tabular}
\end{center}
\caption{\label{tab:res_wsj}
Detailed recognition results in \% WER for the '92 and '93 test sets using 
different acoustic models on the WSJ data.
}
\end{table}

The results on the different test sets on the WSJ data are displayed in
Tab.~\ref{tab:res_rm}.
% TODO Korbinian: The following is for the si-84/half results, as currently
%                 in the table.
Using the parameters calibrated on the RM test data, the {\em 2lvl} system
performance is not as good as of the continuous or GMM systems, but still
in a similar range. Once the number of Gaussians and leaves, as well as the
smoothing parameters are tuned on the development set, we expect the error
rates to be in between the continuous and SGMM systems, as seen on the RM 
data.

\section{Summary}
In this article, we compared continuous models and SGMM to 
two types of semi-continuous hidden Markov models, one using a single codebook 
and the other using multiple codebooks based on a two-level phonetic decision tree.
%
While the first could not produce convincing results, the multiple-codebook 
architecture shows promising performance, especially for limited training data.

Although the current performance is below the state-of-the-art, the
rather simple theory and low computational complexity, paired with the possibility
of solely acoustic adaptation make two-level tree based semi-continuous acoustic
models an attractive alternative to low-resource applications -- both in terms
of computational power and training data.
%
The take home message from the experiments is that the choice of acoustic model
should be made based on a resource constraint (number of Gaussians, available
compute power), and that both continuous and semi-continuous models can do the job.

@@

% Korbinian: We might need that ;)
%\footnotesize
\bibliographystyle{IEEEbib}
\bibliography{refs-eig,refs}

>>>>

\begin{table}%[tb]
\begin{center}
%\footnotesize
%\begin{tabular}{|l||r|r|r|r|r|r||c|}
\begin{tabular}{|l||c|c|c|c|c|c||c|}
\hline
~               & ma87 & oc87 & fe89 & oc89 & fe91 & se92 & {\em avg}  \\ \hline\hline
%{\em mono/cont} &  6.24 &  8.72 &  7.85 &  9.80 &  8.21 & 13.60 & 9.50 \\ \hline
%{\em tri1/cont} &  0.96 &  2.76 &  2.69 &  3.61 &  3.30 &  6.33 & 3.64 \\ \hline\hline
{\em cont} &  1.08 &  2.48 &  2.69 &  3.46 &  2.66 &  5.90 & 3.38 \\ \hline
{\em semi} &  1.80 &  3.19 &  4.72 &  4.62 &  4.15 &  6.88 & 4.66 \\ \hline 
{\em 2lvl} &  0.48 &  1.70 &  2.46 &  3.35 &  1.89 &  5.31 & 2.90 \\ \hline
{\em sgmm} &  0.48 &  2.20 &  2.62 &  2.50 &  1.93 &  5.12 & 2.78 \\ \hline
\end{tabular}
\end{center}
\caption{\label{tab:res_rm}
Detailed recognition results in \% WER for the six DARPA test sets using 
different acoustic models on the RM data set.
}
\end{table}

@@

% TODO Korbinian: More expts are ongoing; these include the full training set
%                 and cmvn, and hopefully a good param combination for 2lvl...
\begin{table}%[tb]
\begin{center}
\begin{tabular}{|l||c|c|c|c|c|c||c|}
\hline
~                & eval '92 & eval '93 & {\em avg} \\ \hline\hline
% all experiments si-84/half, cmvn
{\em cont}       &    12.75 &    17.10 &     14.39 \\ \hline
{\em 2lvl/none}  &    12.92 &    17.85 &     14.79 \\ \hline
{\em 2lvl/intra} &    13.03 &    17.56 &     14.61 \\ \hline
{\em 2lvl/inter} &    12.80 &    17.59 &     14.61 \\ \hline
{\em 2lvl/both}  &    13.01 &    17.59 &     14.75 \\ \hline
{\em sgmm}       &    11.72 &    14.25 &     12.68 \\ \hline
\end{tabular}
\end{center}
\caption{\label{tab:res_wsj}
Detailed recognition results in \% WER for the '92 and '93 test sets using 
different acoustic models on the WSJ data; the {\em 2lvl} system was trained
with and without the different interpolations.
}
\end{table}

The results on the different test sets on the WSJ data are displayed in
Tab.~\ref{tab:res_rm}.
% TODO Korbinian: The following is for the si-84/half results, as currently
%                 in the table.
Using the parameters calibrated on the RM test data, the {\em 2lvl} system
performance is not as good as of the continuous or GMM systems, but still
in a similar range while using about the same number of leaves but about a
third of the Gaussians.
Once the number of Gaussians and leaves, as well as the smoothing parameters 
are tuned on the development set, we expect the error rates to be in between 
the continuous and SGMM systems, as seen on the RM data.

\section{Summary}
In this article, we compared continuous and SGMM models to 
two types of semi-continuous hidden Markov models, one using a single codebook 
and the other using multiple codebooks based on a two-level phonetic decision tree.
%
While the first could not produce convincing results, the multiple-codebook 
architecture shows promising performance, especially for limited training data. 
%
% TODO This sentence only for SI-84/half experiments
The experiments on the WSJ data need to be extended to the full training set, and
the parameter settings calibrated on the development set.

Although the current performance is below the state-of-the-art, the
rather simple theory and low computational complexity, paired with the possibility
of solely acoustic adaptation make multiple-codebook semi-continuous acoustic
models an attractive alternative for low-resource applications -- both in terms
of computational power and training data; especially for embedded applications,
where the codebook evaluation can be implemented in hardware, thus eliminating
the time difference in evaluating full and diagonal covariance Gaussians.
%
The take home message from the experiments is that the choice of acoustic model
should be made based on a resource constraint (number of Gaussians, available
compute power), and that both continuous and semi-continuous models can do the job.

@@

% Korbinian: We might need that ;)
\footnotesize
\bibliographystyle{IEEEbib}
\bibliography{refs-eig,refs}
